WrapITK: Enhanced languages support\\ for the Insight Toolkit

### 2006-07-27 article/Article.tex
grammar and spelling fixes from Michael Bell
darcs-hash:20060727184855-2fc9d-321d2376a056a86b8739214fdef788aa27b7963f.gz

\begin{abstract}
\noindent
ITK \cite{ITKWebSite} is a huge image analysis library, which contains lots of state of the arts
algorithms implementations. However, using it in C++ can be difficult and is
definitively bad suited for prototyping. WrapITK aims to allow classes from ITK
(and custom, classes that interact with ITK) to be "wrapped" for use with
languages like Python \cite{PythonWebSite}, Tcl \cite{TclWebSite}, and Java \cite{JavaWebSite}.
\end{abstract}

@@

WrapITK is a project designed to allow classes from ITK (and custom, classes
that interact with ITK) to be "wrapped" for use with languages like
Python \cite{PythonWebSite}, Tcl \cite{TclWebSite}, and Java \cite{JavaWebSite}.

Note that ITK already has a wrapping infrastructure, and that WrapITK is based
on it, and use the same tools: CMake \cite{CMakeWebSite}, GCC-XML \cite{GccxmlWebsite}
and CableSwig\footnote{CableSwig is based on a now quite old version of SWIG
\cite{SwigWebSite}.} \cite{CableSwigWebSite}. This project aims to
address the following deficits of the existing wrappers (and others):
\begin{itemize}
  \item  ITK is a huge library, but only a small number of classes are available
in target languages. It become quickly frustrating for the user, especially when
he has to spend lot of time to extend the current set of classes. Even if it is
not yet complete, the WrapITK's set of classes have been highly extended. Moreover,
the user can choose at build time which types and which dimensions he want to wrap.
With 202 wrapped filters, WrapITK covers 63\% of the available filters.

  \item  The template arguments set is poorly chosen, making sometime impossible
to create a pipeline. In WrapITK, most of the filters have the same input and output
types, and only a few filters allow to change type. This make the types manipulated 
by filters more consistent, and the user should always be able to build his pipeline.

  \item  Lots of types returned by ITK object's methods are not usable in target
languages. For example, the \verb$GetPixel()$ method of the class \verb$itk::Image$
returns a string describing a pointer, but don't return the pixel value. In WrapITK,
most types used in classes are available in target languages.

  \item  Names in target languages are inconsistent. WrapITK use a strict naming
convention which should make easier to identify the template arguments.

@@

  \item  It is difficult if not impossible to add original-style ITK wrappers for
external C++ classes that interact with ITK. WrapITK provides explicit hooks for
external C++ classes to be wrapped and even installed in the WrapITK tree so
that they interact seamlessly with the other wrapped classes.

  \item  The python's \verb$InsightToolkit$ module is only structured as a big
list of names. It make it nearly unusable in the python interpreter. WrapITK
comes with a new well designed python module easy to use in interpreter, and
providing run-time lookup of templated types - thing which can't be easily done
in C++. Additionally, WrapITK ensures that \verb$SmartPointer$s are always
returned and acceptable as input, so no bare pointers are ever exposed to
Python. This is not the case in the standard ITK wrappers.

@@

  \item  Loading python modules can take lot of time. With WrapITK, by default,
only the modules really used are loaded, and so the loading time is higly shorter
in common situations.

@@

The article you're reading is not as nice and complete than what we would have
done, but it seems important to us to release our work and to get feedbacks
as soon as possible. The article will continue to evolve with WrapITK.

@@

Java, Tcl and Python builds properly and are fully usable. However, Java and
Tcl don't yet have the extended features added to Python\footnote{Any help
to extend Java and Tcl support would be higly appreciated.}.

WrapITK is mainly developed on Mandriva Linux \cite{MandrivaWebSite} and 
MacOS X \cite{MacOsXWebSite} and so is well tested on this palteforms.
It also builds on windows \cite{WindowsWebSite}, but require more testing
to be sure everything work as it should. See below for details about how
to report bugs are contribute patches.

@@

WrapITK provides an interface to some C++ compiled code, and so, execution time
are very similar to pure C++ programs in most of cases.

The major difference comes from the memory
usage: while a C++ program will produce a binary executable containing only
the required code, WrapITK binary contains all the classes wrapped, and so
WrapITK can take a significant amount of memory.

Some comvenient features, like sequence management in python, can be quite
inefficient and should not be used in a loop.

@@

The last version can be obtained from the development repository with darcs \cite{DarcsWebSite}.
The command is
\verb$darcs get --partial http://voxel.jouy.inra.fr/darcs/contrib-itk/WrapITK/$\footnote{Note
that the {\em --partial} option is required on systems with case insensitive filesystem like
windows or Mac Os X.}.

For the user who don't want to use darcs \cite{DarcsWebSite} but still want the last development version,
a nightly updated archive is available at
\url{http://voxel.jouy.inra.fr/darcs/contrib-itk/WrapITK/WrapITK.tar.gz} or
\url{http://voxel.jouy.inra.fr/darcs/contrib-itk/WrapITK/WrapITK.zip}.

@@

Some required patch may appear in the development version version of WrapITK. Those
patches are required for the last stable version of ITK, and should be already
integrated in the last CVS version of ITK.

@@

Python \cite{PythonWebSite} is required only to build python support.
WrapITK is reported to work with python 2.3 and python 2.4. However, the test framework
requires the \verb$subprocess$ module available in standard only in python 2.4 and above.
To run the python test with python 2.3, you have to install \verb$subprocess$.

@@

Tcl \cite{TclWebSite} is required only to build tcl support.
WrapITK have been tested with Tcl 8.4.11.

@@

Java \cite{JavaWebSite} is required only to build java support.
WrapITK have been tested with java \verb$1.5.0_06-b05$ and \verb$1.4.2_11$.

@@

The project is provided with defaults build option which should OK for most of
users. However, for specific needs, you might want to change those options:
\begin{itemize}
  \item \verb$WRAP_TEMPLATE_IF_DIMS$ is the list of dimensions which will be available in the
target languages. The dimensions must be separated by a semicolon (\verb$;$).
By default dimensions 2 and 3 are available.
  \item \verb$WRAP_covariant_vector_double$, \verb$OFF$ by default.
  \item \verb$WRAP_covariant_vector_float$, \verb$ON$ by default.
  \item \verb$WRAP_double$ \verb$OFF$, by default.
  \item \verb$WRAP_float$ \verb$ON$, by default. Note that it is the only signed
type selected by default, so you will have to use floats to manipulate signed
values.
  \item \verb$WRAP_rgb_unsigned_char$, \verb$ON$ by default.
  \item \verb$WRAP_rgb_unsigned_short$, \verb$OFF$ by default.
  \item \verb$WRAP_signed_char$, \verb$OFF$ by default.
  \item \verb$WRAP_signed_long$, \verb$OFF$ by default.
  \item \verb$WRAP_signed_short$, \verb$OFF$ by default.
  \item \verb$WRAP_unsigned_char$, \verb$OFF$ by default.
  \item \verb$WRAP_unsigned_long$, \verb$OFF$ by default. Some filters, like
\verb$WatershedImageFilter$ require this type. Some filters to return to a
wrapped type from \verb$unsigned long$ are provided, even if this option is
set to \verb$OFF$.
  \item \verb$WRAP_unsigned_short$, \verb$ON$ by default. \verb$unsigned short$
is the only integer type available by default. This type have been choose rather
than \verb$unsigned char$ to be able to manipulate 8-bits as well as 16-bits images,
and to be able to manipulate labeled images more than 255 labels. It is still
possible to save images with the \verb$unsigned char$ type, even if
\verb$WRAP_unsigned_char$ is set to \verb$OFF$.
  \item \verb$WRAP_vector_double$, \verb$OFF$ by default.
  \item \verb$WRAP_vector_float$, \verb$ON$ by default.

@@

The user should modify those options carefully: activate all the types, and/or
add lots of dimensions will produce very large binary files which will take lots
of memory once loaded.

Note that each individual filter that is wrapped can declare which dimensions it
should be wrapped for, and what image types it can accept. For example, a filter
could declare that it should only be wrapped for 3D images with floating-point
typed pixels. In this case, then wrappers will only be created if the user has
selected to build 3-dimensional image wrappers and has selected one or more
floating point types (e.g. double or float) in ccmake. Thus, the ccmake
configuration specifies the maximum possible range of image and filter types to
be created, and each filter is wrapped for some subset of that range.

Project should always be built outside the source directory, in a \verb$build$
directory for example.

@@

RPM packages for Mandriva Linux 2006 are available at \url{http://voxel.jouy.inra.fr/mdk/mima2}.
To install WrapITK for mandriva linux 2006.0, just add a new media with the
command

@@

This media also contain several packages which may be useful to use WrapITK,
and which are not available in mandriva linux 2006.0. Here is the full list:
\begin{itemize}
  \item \verb$cmake$
  \item \verb$darcs$
  \item \verb$itk-data$
  \item \verb$itk-doc$
  \item \verb$itk-examples$
  \item \verb$itkvtk-devel$
  \item \verb$libitk$
  \item \verb$libitk-devel$
  \item \verb$libvtk$
  \item \verb$libvtk-devel$
  \item \verb$libvtk-qt$
  \item \verb$python-itk$
  \item \verb$python-itk-numarray$
  \item \verb$python-itkvtk$
  \item \verb$python-vtk$
  \item \verb$tcl-itk$
  \item \verb$tcl-vtk$
  \item \verb$vtk-data$
  \item \verb$vtk-doc$
  \item \verb$vtk-examples$
  \item \verb$vtk-test-suite$
  \item \verb$wrapitk-devel$
\end{itemize}

@@

In this section, we detail python usage. Some of the examples shown here are copied
from the console, and so show the interpreter prompt:

@@

In the example above, \verb$12+3$ is what is wrote in the interpreter, and
\verb$15$ is the result. \verb$2>$, \verb$3>$, \verb$4>$ are the prompt
of the interpreter.

@@

On linux boxes however, the user have to set the \verb$LD_LIBRARY_PATH$ to
point to libSwigRuntime.so. For example
\verb$export LD_LIBRARY_PATH=/usr/lib/InsightToolkit/WrapITK/Python-SWIG$.
This step is not required with the mandriva's package.

@@

image = imageType.New()
\end{verbatim}
This construction is similar to what is done in C++, and make it easy to change
the dimension used for example - it can even be changed at run-time.

A more convenient syntax for usage in interpreter is also available:
\begin{itemize}
  \item \verb$itk.Image.UC3$
  \item \verb$itk.Image.US3$
\end{itemize}
\verb$itk.Image.UC3$ refere to the same class than \verb$itk.Image[itk.UC, 3]$
but have the advantage to allow to use the tab-completion in the interpreter,
and so let the user easily know which template arguments he can use.
However, this notation is more rigid than the one above and won't let the user
specify the type and the dimension used in a single place, and thus,
should be use only in interpreter.

Filters templated on images can be similarly accessed:
\begin{itemize}
  \item \verb$itk.ImageFileReader[itk.Image[itk.UC,3]]$
  \item or \verb$itk.ImageFileReader[itk.Image.UC3]$
  \item or \verb$itk.ImageFileReader.IUC3$
  \item or \verb$itk.ImageFileReader[imageType]$
  \item or even with an instance of the class used as template parameter: \verb$itk.ImageFileReader[image]$.
\end{itemize}
This makes it easy to write generic routines which
can deal with any input image type. For example, a function which take an image
as parameter and write it to a file without having to give the image type can be:
\begin{verbatim}
def write( image, fileName ) :
    writer = itk.ImageFileWriter[ image ].New()
    writer.SetFileName( fileName )
    writer.SetInput( image )
    writer.Update()
\end{verbatim}

@@

Lots of classes have a \verb$New()$ method which returns a smart pointer to an object of
that class. In python, the \verb$New()$ method has some additional features:
\begin{itemize}
  \item Arguments to the new method are assumed to be filter inputs. So you could
write:
\begin{verbatim}
adder = itk.AddImageFilter[...].New()
adder.SetInput1( readerA.GetOutput() )
adder.SetInput2( readerB.GetOutput() )
\end{verbatim}
or you could write
\begin{verbatim}
adder = itk.AddImageFilter[...].New( readerA.GetOutput(), readerB.GetOutput() )
\end{verbatim}
or even
\begin{verbatim}
adder = itk.AddImageFilter[...].New( readerA, readerB )
\end{verbatim}
In that case, \verb$New()$ will use the \verb$GetOutput()$ method of the object, if it
exist, to get the image and set the inputs of the new filter.

  \item Additionally, keyword arguments are allowed as well. Keyword arguments cause
the corresponding \verb$Set...$ method to be called, so you could write the
following:
\begin{verbatim}
itk.ImageFileWriter[image].New(image, FileName="foo.tif")
\end{verbatim}
or
\begin{verbatim}
itk.ImageFileWriter[image].New(Input=image, FileName="foo.tif")
\end{verbatim}.

@@

and, more important, most of classes can be instantiated and parametered in one line,
which make ITK less verbose, and a lots more easy to use in the interpreter.

@@

To set the radius of a \verb$MedianImageFilter$ object, for example, the user have to
create a \verb$Size$ object and use it as argument of the \verb$SetRadius()$ method.
\begin{verbatim}
12> radius = itk.Size[2]()

@@

Note that the \verb$SetElement()$ method doesn't check the bound of the object, and thus
is unsafe. The following code is executed, and can lead to a segmentation fault.

@@

This time, a bound check is performed, and the user is not able to use an invalid index.

@@

Even if it's a safe method, it is still not really convenient.

@@

Also, with the optional patches, some itk object can be converted to python sequences.

@@

Some convenient functions are provided with the itk module. They all begin with a lower
case character to clearly show they are not part of ITK.
\begin{itemize}
  \item \verb$itk.image(object)$ try to return an image from the object given in
parameter. If the object is an image, it is returned without changes. If the object
is a filter, the object returned by \verb$GetOutput()$ method is returned.
This function is used in most of the next functions to allow the user to
pass an image or a filter, and is available here for the same usage in some custom
fuctions.

  \item \verb$itk.range(object)$ return the range of values of an image in a tuple.
\verb$object$ can be an image or a filter. In case of a filter, \verb$itk.image()$ is used
to get the output image of the filter. The function update the pipeline by calling
\verb$UpdateOutputInformation()$ and \verb$Update()$.

This function is only a convenient function for a common task while prototyping.

@@

  \item \verb$itk.size(object)$ return the size of an image.
\verb$object$ can be an image or a filter. In case of a filter, \verb$itk.image()$ is used
to get the output image of the filter. The function update only the information of the
pipeline, by calling \verb$UpdateOutputInformation()$, but don't trigger a full update
of the pipeline.

This function is only a convenient function for a common task while prototyping.

@@

  \item \verb$itk.write(object, fileName)$ write an image in a file, without
having to pass the image type.
\verb$object$ can be an image or a filter. In case of a filter, \verb$itk.image()$ is used
to get the output image of the filter. The function update the pipeline by calling
\verb$UpdateOutputInformation()$ and \verb$Update()$.

This function is only a convenient function for a common task while prototyping.

@@

  \item \verb$itk.show()$, \verb$itk.show2D()$ and \verb$itk.show3D()$ are used to
display images. \verb$itk.show2D()$ requires to have imview \cite{ImviewWebSite} installed.
\verb$itk.show3D()$ requires to have Vtk for python, ItkVtkGlue\footnote{ItkVtkGlue can be found in
the {\em ExternalProject} directory of WrapITK} for python, PyQt \cite{PyQtWebSite}, and iPython
\cite{IPythonWebSite} installed, and to use iPython with the \verb$-qthread$ option.

@@

\verb$itk.show3D()$ display a volumic rendering of the image. See Figure~\ref{screenshot}.

@@

  \item \verb$itk.strel(d, s)$ is used to create a binary ball structuring of dimension
\verb$d$ and size \verb$s$. Structuring element support is quite bad currently in WrapITK
and should change in the future. Using \verb$itk.strel$ rather than creating a
\verb$BinaryBallStructuringElement$ directly is recommended to have backward compatibility
when the structuring element type will be changed.

@@

\verb$itk.auto_progress(True)$ also set an import callback which show the module name when
the module are imported.

  \item \verb$itk.class_(object)$ return the class of an object. The \verb$__class__$
attribute is often not what the user want with ITK. \verb$itk.class_$ is a convenient
function to get the class of an ITK object.

@@

  \item \verb$itk.echo(object, file)$ is a convenient function to call the \verb$Print()$
method of an ITK object without the need to pass a \verb$StringStream$ object.
This function is less useful with the optional patches: the \verb$__str__()$
method do a very similar job with a better integration with python.

@@

  \item \verb$itk.pipeline$ class let the developer easily create
a custom pipeline which can then be manipulated as a pure ITK filter.
It provide several methods:
\begin{itemize}
  \item \verb$__init__( self, input=None )$ is the constructor of the pipeline.
The input of the pipeline can be passed as parameter.

  \item \verb$connect( self, filter )$ connect a new filter to the pipeline.
The output of the last filter in the pipeline will be set as the input of the filter
passed as parameter, and the filter passed as parameter will be added to the filter
list.

@@

  \item \verb$GetInput( self )$ return the input of the last filter in the pipeline.
If another input is needed, use \verb$pipeline[0].GetAnotherInput()$ instead of
this method, or subclass pipeline to implement another \verb$GetInput()$ method.

@@

  \item \verb$__getitem__( self, i )$ and \verb$__len__( self )$ provide common
python list manipulation interface to the pipeline object.

@@

Example:
ITK current implementation of morphological dilation, erosion, opening and closing
can be very inefficient with large structuring elements. Also, WrapITK only give
access to ball structuring element. The following class illustrate the use of
\verb$itk.pipeline$ class to implement a efficient opening in 3 dimensions with
a box structuring element. We here take advantage of structuring element decomposition:
a dilation (or erosion) by a box can be more efficiently computed by perfoming
3 dilations (or erosion) with line structuring element oriented on each dimension.
The \verb$itk.pipeline$ encapsulate the 6 filters needed to perform the efficient
opening, and take care of setting the structuring element for all the internal
filters in the \verb$SetKernel(self, x,y,z)$, according to the size wanted by the
user.

@@

This script is the exact transcription to python of the C++ example 
which can be found at Examples/Filtering/GradientMagnitudeRecursiveGaussianImageFilter.cxx
in the ITK source tree. More information about the filters used can be found in 
the ITK Software Guide \cite{{ITKSoftwareGuide}}, section 6.4.2.

@@

  \item \verb$darcs whatsnew$ to display the changes you have done in
your copy of the repository.

>>>>

\begin{abstract}
\noindent
ITK \cite{ITKWebSite} is a huge image analysis library, which contains lots of state of the art
algorithms implementations. However, using it in C++ can be difficult and is
poorly suited for prototyping. WrapITK aims to allow classes from ITK
(and custom, classes that interact with ITK) to be "wrapped" for use with
languages like Python \cite{PythonWebSite}, Tcl \cite{TclWebSite}, and Java \cite{JavaWebSite}.
\end{abstract}

@@

WrapITK is a project designed to allow classes from ITK (and custom classes
that interact with ITK) to be "wrapped" for use with languages like
Python \cite{PythonWebSite}, Tcl \cite{TclWebSite}, and Java \cite{JavaWebSite}.

Note that ITK already has a wrapping infrastructure, and that WrapITK is based
on it, and use the same tools: CMake \cite{CMakeWebSite}, GCC-XML \cite{GccxmlWebsite}
and CableSwig\footnote{CableSwig is based on a now quite old version of SWIG
\cite{SwigWebSite}.} \cite{CableSwigWebSite}. This project aims to
address the following deficits of the existing wrappers (and others):
\begin{itemize}
  \item  ITK is a huge library, but only a small number of classes are available
in target languages. This quickly becomes frustrating for the user, especially when
he has to spend lot of time to extend the current set of classes. Even if it is
not yet complete, the WrapITK's set of classes have been highly extended. Moreover,
the user can choose at build time which types and which dimensions he wants to wrap.
With 202 wrapped filters, WrapITK covers 63\% of the available filters.

  \item  The template argument set is poorly chosen, sometimes making it impossible
to create a pipeline. In WrapITK, most of the filters have the same input and output
types, and only a few filters are allowed to change type. This make the types manipulated 
by filters more consistent, and the user should always be able to build his pipeline.

  \item  Many types returned by ITK object's methods are not usable in target
languages. For example, the \verb$GetPixel()$ method of the class \verb$itk::Image$
returns a string describing a pointer, but does not return the pixel value. In WrapITK,
most types used in the classes are available in target languages.

  \item  Names in target languages are inconsistent. WrapITK uses a strict naming
convention which should make it easier to identify the template arguments.

@@

  \item  It is difficult if not impossible to add original-style ITK wrappers for
external C++ classes that interact with ITK. WrapITK provides explicit hooks for
external C++ classes to be wrapped and even installed in the WrapITK tree so
they interact seamlessly with the other wrapped classes.

  \item  The python's \verb$InsightToolkit$ module is only structured as a big
list of names which makes it nearly unusable in the python interpreter. WrapITK
comes with a new, well-designed python module that is easy to use with the interpreter, and
which provides run-time lookup of templated types - things which can't be easily done
in C++. Additionally, WrapITK ensures that \verb$SmartPointer$s are always
returned and acceptable as input, so no bare pointers are ever exposed to
Python. This is not the case in the standard ITK wrappers.

@@

  \item  Loading python modules can take lot of time. With WrapITK, by default,
only the modules really used are loaded, so the loading time is much shorter
in common situations.

@@

The article you're reading is not yet complete,
but it seems important for us to release our work and to get feedback
as soon as possible. The article will continue to evolve with WrapITK.

@@

Java, Tcl and Python build properly and are fully usable. However, Java and
Tcl don't yet have the extended features added to Python\footnote{Any help
to extend Java and Tcl support would be higly appreciated.}.

WrapITK is mainly developed on Mandriva Linux \cite{MandrivaWebSite} and 
MacOS X \cite{MacOsXWebSite} and therefore is well tested on these platforms.
It also builds on windows \cite{WindowsWebSite}, but more testing is required
to be sure everything works as it should. See below for details about how
to report bugs and contribute patches.

@@

WrapITK provides an interface to some C++ compiled code, so execution times
are very similar to pure C++ programs in most of cases.

The major difference comes from the memory
usage: while a C++ program will produce a binary executable containing only
the required code, the WrapITK binary contains all the wrapped classes. So,
WrapITK can take a significant amount of memory.

Some convenient features, like sequence management in python, can be quite
inefficient and should not be used in a loop.

@@

The latest version can be obtained from the development repository with darcs \cite{DarcsWebSite}.
The command is
\verb$darcs get --partial http://voxel.jouy.inra.fr/darcs/contrib-itk/WrapITK/$\footnote{Note
that the {\em --partial} option is required on systems with case insensitive filesystems like
windows or Mac OS X.}.

For the user who does not want to use darcs \cite{DarcsWebSite} but still wants the last development version,
a nightly updated archive is available at
\url{http://voxel.jouy.inra.fr/darcs/contrib-itk/WrapITK/WrapITK.tar.gz} or
\url{http://voxel.jouy.inra.fr/darcs/contrib-itk/WrapITK/WrapITK.zip}.

@@

Some required patches may appear in the development version of WrapITK. Those
patches are required for the last stable version of ITK, and should be already
integrated in the last CVS version of ITK.

@@

Python \cite{PythonWebSite} is required only to build python support.
WrapITK is reported to work with python 2.3 and python 2.4. However, the test framework
requires the \verb$subprocess$ module to be available, which is standard only in python 2.4 and above.
To run the python test with python 2.3, you have to install \verb$subprocess$.

@@

Tcl \cite{TclWebSite} is required only to build tcl support.
WrapITK has been tested with Tcl 8.4.11.

@@

Java \cite{JavaWebSite} is required only to build java support.
WrapITK has been tested with java \verb$1.5.0_06-b05$ and \verb$1.4.2_11$.

@@

The project is provided with default build options which should be OK for most
users. However, for specific needs, you might want to change these options:
\begin{itemize}
  \item \verb$WRAP_TEMPLATE_IF_DIMS$ is the list of dimensions which will be available in the
target languages. The dimensions must be separated by a semicolon (\verb$;$).
By default dimensions 2 and 3 are available.
  \item \verb$WRAP_covariant_vector_double$, \verb$OFF$ by default.
  \item \verb$WRAP_covariant_vector_float$, \verb$ON$ by default.
  \item \verb$WRAP_double$ \verb$OFF$, by default.
  \item \verb$WRAP_float$ \verb$ON$, by default. Note that float is the only signed
type selected by default, so you will have to use floats to manipulate signed
values.
  \item \verb$WRAP_rgb_unsigned_char$, \verb$ON$ by default.
  \item \verb$WRAP_rgb_unsigned_short$, \verb$OFF$ by default.
  \item \verb$WRAP_signed_char$, \verb$OFF$ by default.
  \item \verb$WRAP_signed_long$, \verb$OFF$ by default.
  \item \verb$WRAP_signed_short$, \verb$OFF$ by default.
  \item \verb$WRAP_unsigned_char$, \verb$OFF$ by default.
  \item \verb$WRAP_unsigned_long$, \verb$OFF$ by default. Some filters, like
\verb$WatershedImageFilter$ require this type. Some filters to return to a
wrapped type from \verb$unsigned long$ are provided, even if this option is
set to \verb$OFF$.
  \item \verb$WRAP_unsigned_short$, \verb$ON$ by default. \verb$unsigned short$
is the only integer type available by default. This type has been choosen rather
than \verb$unsigned char$ to be able to manipulate 8-bits as well as 16-bits images,
and to be able to manipulate labeled images more than 255 labels. It is still
possible to save images with the \verb$unsigned char$ type, even if
\verb$WRAP_unsigned_char$ is set to \verb$OFF$.
  \item \verb$WRAP_vector_double$, \verb$OFF$ by default.
  \item \verb$WRAP_vector_float$, \verb$ON$ by default.

@@

The user should modify these options carefully: activate all the types, and/or
adding many dimensions will produce very large binary files which will take a lot
of memory once loaded.

Note that each individual filter that is wrapped can declare which dimensions it
should be wrapped for, and what image types it can accept. For example, a filter
could declare that it should only be wrapped for 3D images with floating-point
typed pixels. In this case, the wrappers will only be created if the user has
chosen to build 3-dimensional image wrappers and has selected one or more
floating point types (e.g. double or float) in ccmake. Thus, the ccmake
configuration specifies the maximum possible range of image and filter types to
be created, and each filter is wrapped for some subset of that range.

Projects should always be built outside the source directory, in a \verb$build$
directory for example.

@@

RPM packages for Mandriva Linux 2006 are available at \url{http://voxel.jouy.inra.fr/mdk/mima2}.
To install WrapITK for Mandriva Linux 2006.0, just add a new media with the
command

@@

This media also contains several packages which may be useful to use WrapITK,
and which are not available in mandriva linux 2006.0. Here is the full list:
\begin{itemize}
  \item \verb$cmake$
  \item \verb$darcs$
  \item \verb$itk-data$
  \item \verb$itk-doc$
  \item \verb$itk-examples$
  \item \verb$itkvtk-devel$
  \item \verb$libitk$
  \item \verb$libitk-devel$
  \item \verb$libvtk$
  \item \verb$libvtk-devel$
  \item \verb$libvtk-qt$
  \item \verb$python-itk$
  \item \verb$python-itk-numarray$
  \item \verb$python-itkvtk$
  \item \verb$python-vtk$
  \item \verb$tcl-itk$
  \item \verb$tcl-vtk$
  \item \verb$vtk-data$
  \item \verb$vtk-doc$
  \item \verb$vtk-examples$
  \item \verb$vtk-test-suite$
  \item \verb$wrapitk-devel$
\end{itemize}

@@

In this section, we detail python usage. Some of the examples shown here are copied
from the console which shows the interpreter prompt:

@@

In the example above, \verb$12+3$ is what is written in the interpreter, and
\verb$15$ is the result. \verb$2>$, \verb$3>$, \verb$4>$ are the prompts
of the interpreter.

@@

On linux boxes however, the user must set the \verb$LD_LIBRARY_PATH$ to
point to libSwigRuntime.so. For example
\verb$export LD_LIBRARY_PATH=/usr/lib/InsightToolkit/WrapITK/Python-SWIG$.
This step is not required with the mandriva's package.

@@

image = imageType.New()
\end{verbatim}
This construction is similar to what is done in C++, and makes it easy to change
the dimension used. For example - it can even be changed at run-time.

A more convenient syntax for usage in the interpreter is also available:
\begin{itemize}
  \item \verb$itk.Image.UC3$
  \item \verb$itk.Image.US3$
\end{itemize}
\verb$itk.Image.UC3$ refers to the same class as \verb$itk.Image[itk.UC, 3]$
but has the advantage of allowing the use of tab-completion in the interpreter,
and lets the user easily know which template arguments he can use.
However, this notation is more rigid than the one above and won't let the user
specify the type and the dimension used in a single place. Therefor, this syntax
should be used only in interpreter.

Filters templated on images can be similarly accessed:
\begin{itemize}
  \item \verb$itk.ImageFileReader[itk.Image[itk.UC,3]]$
  \item or \verb$itk.ImageFileReader[itk.Image.UC3]$
  \item or \verb$itk.ImageFileReader.IUC3$
  \item or \verb$itk.ImageFileReader[imageType]$
  \item or even with an instance of the class used as a template parameter: \verb$itk.ImageFileReader[image]$.
\end{itemize}
This makes it easy to write generic routines which
can deal with any input image type. For example, a function which takes an image
as parameter and writes it to a file without having to give the image type can be:
\begin{verbatim}
def write( image, fileName ) :
    writer = itk.ImageFileWriter[ image ].New()
    writer.SetFileName( fileName )
    writer.SetInput( image )
    writer.Update()
\end{verbatim}

@@

Many classes have a \verb$New()$ method which returns a smart pointer to an object of
that class. In python, the \verb$New()$ method has some additional features:
\begin{itemize}
  \item Arguments to the new method are assumed to be filter inputs. So you could
write:
\begin{verbatim}
adder = itk.AddImageFilter[...].New()
adder.SetInput1( readerA.GetOutput() )
adder.SetInput2( readerB.GetOutput() )
\end{verbatim}
or you could write
\begin{verbatim}
adder = itk.AddImageFilter[...].New( readerA.GetOutput(), readerB.GetOutput() )
\end{verbatim}
or even
\begin{verbatim}
adder = itk.AddImageFilter[...].New( readerA, readerB )
\end{verbatim}
In that case, \verb$New()$ will use the \verb$GetOutput()$ method of the object, if it
exists, to get the image and set the inputs of the new filter.

  \item Additionally, keyword arguments are allowed . Keyword arguments cause
the corresponding \verb$Set...$ method to be called, so you could write the
following:
\begin{verbatim}
itk.ImageFileWriter[image].New(image, FileName="foo.tif")
\end{verbatim}
or
\begin{verbatim}
itk.ImageFileWriter[image].New(Input=image, FileName="foo.tif")
\end{verbatim}.

@@

and, more importantly, most of the classes can be instantiated and parameterized in one line,
which make ITK less verbose, and a lot more easy to use in the interpreter.

@@

To set the radius of a \verb$MedianImageFilter$ object, for example, the user has to
create a \verb$Size$ object and use it as an argument of the \verb$SetRadius()$ method.
\begin{verbatim}
12> radius = itk.Size[2]()

@@

Note that the \verb$SetElement()$ method doesn't check the bounds of the object, and thus
is unsafe. The following code is executed, and can lead to a segmentation fault.

@@

This time, a bounds check is performed, and the user is not able to use an invalid index.

@@

Even if it is a safe method, it is still not really convenient.

@@

Also, with the optional patches, some itk objects can be converted to python sequences.

@@

Some convenience functions are provided with the itk module. They all begin with a lower
case character to clearly show they are not part of ITK.
\begin{itemize}
  \item \verb$itk.image(object)$ try to return an image from the object given in
parameter. If the object is an image, it is returned without changes. If the object
is a filter, the object returned by \verb$GetOutput()$ method is returned.
This function is used in most of the following functions to allow the user to
pass an image or a filter, and is available here for the same usage in some custom
fuctions.

  \item \verb$itk.range(object)$ returns the range of values of an image in a tuple.
\verb$object$ can be an image or a filter. In case of a filter, \verb$itk.image()$ is used
to get the output image of the filter. The function updates the pipeline by calling
\verb$UpdateOutputInformation()$ and \verb$Update()$.

This function is only a convenience function for a common task while prototyping.

@@

  \item \verb$itk.size(object)$ return the size of an image.
\verb$object$ can be an image or a filter. In the case of a filter, \verb$itk.image()$ is used
to get the output image of the filter. The function updates only the information of the
pipeline, by calling \verb$UpdateOutputInformation()$, but does not trigger a full update
of the pipeline.

This function is only a convenience function for a common task while prototyping.

@@

  \item \verb$itk.write(object, fileName)$ write an image to a file, without
having to pass the image type.
\verb$object$ can be an image or a filter. In the case of a filter, \verb$itk.image()$ is used
to get the output image of the filter. The function updates the pipeline by calling
\verb$UpdateOutputInformation()$ and \verb$Update()$.

This function is only a convenience function for a common task while prototyping.

@@

  \item \verb$itk.show()$, \verb$itk.show2D()$ and \verb$itk.show3D()$ are used to
display images. \verb$itk.show2D()$ requires that imview \cite{ImviewWebSite} be installed.
\verb$itk.show3D()$ requires Vtk for python, ItkVtkGlue\footnote{ItkVtkGlue can be found in
the {\em ExternalProject} directory of WrapITK} for python, PyQt \cite{PyQtWebSite}, and iPython
\cite{IPythonWebSite} with the \verb$-qthread$ option.

@@

\verb$itk.show3D()$ display a volumetric rendering of the image. See Figure~\ref{screenshot}.

@@

  \item \verb$itk.strel(d, s)$ is used to create a binary ball structuring of dimension
\verb$d$ and size \verb$s$. Structuring element support is quite bad currently in WrapITK
and should change in the future. Using \verb$itk.strel$ rather than creating a
\verb$BinaryBallStructuringElement$ directly is recommended to have backward compatibility
when the structuring element type changes.

@@

\verb$itk.auto_progress(True)$ also sets an import callback which show the module name when
the module are imported.

  \item \verb$itk.class_(object)$ returns the class of an object. The \verb$__class__$
attribute is often not what the user wants with ITK. \verb$itk.class_$ is a convenience
function to get the class of an ITK object.

@@

  \item \verb$itk.echo(object, file)$ is a convenience function to call the \verb$Print()$
method of an ITK object without the need to pass a \verb$StringStream$ object.
This function is less useful with the optional patches: the \verb$__str__()$
method does a very similar job with better integration with python.

@@

  \item \verb$itk.pipeline$ class let the developer easily create
a custom pipeline which can then be manipulated as a pure ITK filter.
It provides several methods:
\begin{itemize}
  \item \verb$__init__( self, input=None )$ is the constructor of the pipeline.
The input of the pipeline can be passed as parameter.

  \item \verb$connect( self, filter )$ connect a new filter to the pipeline.
The output of the last filter in the pipeline will be set as the input of the filter
passed as a parameter, and the filter passed as a parameter will be added to the filter
list.

@@

  \item \verb$GetInput( self )$ return the input of the last filter in the pipeline.
If another input is needed, use \verb$pipeline[0].GetAnotherInput()$ instead of
this method, or subclass the pipeline to implement another \verb$GetInput()$ method.

@@

  \item \verb$__getitem__( self, i )$ and \verb$__len__( self )$ provide the common
python list manipulation interface to the pipeline object.

@@

Example:
ITK's current implementation of morphological dilation, erosion, opening and closing
can be very inefficient with large structuring elements. Also, WrapITK only gives
access to ball structuring element. The following class illustrates the use of the 
\verb$itk.pipeline$ class to implement an efficient opening in 3 dimensions with
a box structuring element. We take advantage of structuring element decomposition:
a dilation (or erosion) by a box can be more efficiently computed by perfoming
3 dilations (or erosion) with line structuring element oriented on each dimension.
The \verb$itk.pipeline$ encapsulates the 6 filters needed to perform the efficient
opening, and takes care of setting the structuring elements for all the internal
filters in the \verb$SetKernel(self, x,y,z)$, according to the size wanted by the
user.

@@

This script is the exact transcription to python of the C++ example 
which can be found at Examples/Filtering/GradientMagnitudeRecursiveGaussianImageFilter.cxx
in the ITK source tree. More information about the filters used can be found in 
the ITK Software Guide \cite{ITKSoftwareGuide}, section 6.4.2.

@@

  \item \verb$darcs whatsnew$ to display the changes you have made in
your copy of the repository.
